---
title: "Households below average income (HBAI) statistics"
author: "NIck Malleson"
date: "3 July 2015"
output: html_document
---

```{r init}
library(ggplot2) 
library(reshape2) # For manipulating data ready for ggplot2 (in particular need 'melt')

```

# Introduction

A measure commonly used to quantify rates of relative poverty.

"Statistics on the number and percentage of people living in low-income households in the UK." [https://www.gov.uk/government/collections/households-below-average-income-hbai--2](https://www.gov.uk/government/collections/households-below-average-income-hbai--2)


# Definitions:

  - **BHC**: Before Housing Costs
  - **AHC**: After Housing Costs 
  - **Relative** low income: the proportion of individuals who have household incomes below 60% of the median in the given year.
  - **Absolute** low income: as above, but using the average in 2010/11 (adjusted for inflation) which allows for comparisons over time.
  
# Trends

Trends in HBAI from 1994/95 available from the [Home Office](https://www.gov.uk/government/statistics/households-below-average-income-19941995-to-20132014). In particular, see the [documentation](https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/437246/households-below-average-income-1994-95-to-2013-14.pdf) and the [Supporting Trend Data (CSV files)](https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/437256/hba-2013-2014-supporting-csv-files.zip). The data show trends broken down in a number of different ways.

## Absolute low income (below 60% of median in 2010/11)

```{r}
# Read the trends data
d <- read.csv('./hba-2013-2014-supporting-csv-files/Chapter 3 - Trends.csv')

# Just take the columns that we want
d <- d[, c("Year","X60ABSBHCP_ALL","X60ABSAHCP_ALL")]

# Make it long for ggplot2 (all variables in the same column with a separate identifier column)
d.long <- melt(d, id.vars="Year", value.name = "Percent")

ggplot(data=d.long, aes(x=Year, y=Percent, group=1)) + 
  geom_line(color="red", linetype="dotted") +
  geom_point(color="red") +
  xlab("Year") + ylab("Percentage") + ggtitle("Percentage of individuals in absolute low income")

```

XXXX - HERE - Fix the graph above and explain what it's showing. Definition of variables:

60ABSBHCP_ALL	Percentage of individuals in absolute low income - BHC
60ABSAHCP_ALL	Percentage of individuals in absolute low income - AHC